---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hbal

<!-- badges: start -->
<!--
[![Lifecycle:
experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
-->
[![Lifecycle:
stable](https://img.shields.io/badge/lifecycle-stable-green.svg)](https://www.tidyverse.org/lifecycle/#stablel)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![downloads: CRAN](https://cranlogs.r-pkg.org/badges/grand-total/hbal)](https://www.datasciencemeta.com/rpackages)
<!-- badges: end -->

**R** package for implementing hierarchically regularized entropy balancing proposed by Xu and Yang (2022). It is suitable for estimating average treatment effect on the treated with binary treatments under strong ignorability. **hbal** is an extension of entropy balancing: it automatically  expands the feature space by including higher-order terms (such as squared and cubic terms and interactions) of covariates and then achieves approximate balance on the expanded features using ridge penalties with a hierarchical structure.

**Examples:** R code used in the [tutorial](https://yiqingxu.org/packages/hbal/articles/tutorial.html) can be downloaded from [here](https://raw.githubusercontent.com/xuyiqing/hbal/main/pkgdown/hbal_examples.R).

**Reference:** Eddie Yang & Yiqing Xu (2021). [Hierarchically Regularized Entropy Balancing](https://www.cambridge.org/core/journals/political-analysis/article/hierarchically-regularized-entropy-balancing/7DF29028EABE0B2465D1E1FF35F13DDB). *Political Analysis*, published online.

***

## Installation

You can install the **hbal** package from CRAN: 
```{r eval=FALSE}
install.packages("hbal") 
```

You can also install the up-to-date development version from Github: 
```{r eval=FALSE}
install.packages("devtools") # if not already installed
devtools::install_github("xuyiqing/hbal")
```

**hbal** depends on the following packages, which will be installed automatically when **hbal** is being installed; you can also install them manually:  
```{r eval=FALSE}
install_all <- function(packages) {
  installed_pkgs <- installed.packages()[, "Package"]
  for (pkg in packages) {
    if (!pkg %in% installed_pkgs) {
      install.packages(pkg)
    }
  }
}

packages <- c("estimatr", "glmnet", "ggplot2", "gridExtra", "gtable",
              "nloptr", "Rcpp", "RcppEigen", "stringr", "generics")
install_all(packages)
```

***

### Notes on installation failures

1. Intel-Mac users who have updated to MacOS BigSur or higher will likely encounter compilation problems. See [here](http://yiqingxu.org/public/BigSurError.pdf) for a potential solution.
2. Windows users please consider upgrading R to 4.3.0 or higher and installing the [latest Rtools](https://cran.r-project.org/bin/windows/Rtools/) to avoid C++17 complier errors when installing fastplm. 
3. For Rcpp, RcppArmadillo and MacOS "-lgfortran" and "-lquadmath" error, click [here]( http://thecoatlessprofessor.com/programming/rcpp-rcpparmadillo-and-os-x-mavericks-lgfortran-and-lquadmath-error/) for details.
4. Installation failure related to OpenMP on MacOS, click [here](http://thecoatlessprofessor.com/programming/openmp-in-r-on-os-x/) for a solution.
5. To fix these issues, try installing gfortran from [here]( https://gcc.gnu.org/wiki/GFortranBinaries#MacOS clang4 R Binaries from https://github.com/coatless/r-macos-clang).

## Report bugs

Please report bugs to **yiqingxu [at] stanford.edu** with your sample code and data file. Much appreciated!

