---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hbal

<!-- badges: start -->
<!-- badges: end -->

hbal performs hierarchically regularized entropy balancing such that the covariate moments of the control group match those of the treatment group. hbal automatically expands the covariate space to include higher order terms and uses cross-validation to select variable penalties for the balancing conditions.

## Installation

You can install the released version of hbal from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("hbal")
```
Or the development version from GitHub:

``` r
# install.packages("devtools")
devtools::install_github("xuyiqing/hbal")
```

## Getting started

```{r, message=FALSE}
library(hbal)
```

hbal provides two main functions:

* `hbal()`, which performs hierarchically regularized entropy balancing.

* `att()`, which calculates the average treatment effect on the treated (ATT) from an `hbalobject` returned by `hbal()`. 

## Example

For examples on how to use the package, see [here](https://github.com/xuyiqing/hbal)

## Reference

For a detailed description of the method see: 
