---
title: "hbal: Hierarchically Regularized Entropy Balancing"
output: 
  rmarkdown::html_document:
    theme: lumen
vignette: >
  %\VignetteIndexEntry{hbal}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(ggplot2)
library(estimatr)
require(ebal)
```

***

This page demonstrates the usage of the **hbal** package. **hbal** is an implementation of the method introduced in [Xu \& Yang (2021)](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3807620), which performs hierarchically regularized entropy balancing such that the covariate moments of the control group match those of the treatment group. **hbal** automatically expands the covariate space to include higher order terms and uses cross-validation to select variable penalties for the balancing conditions.

**hbal** provides two main functions:

* `hbal`, which performs hierarchically regularized entropy balancing.

* `att`, which calculates the average treatment effect on the treated (ATT) from an `hbalobject` returned by `hbal`. 

***

**Authors:** [Yiqing Xu](http://yiqingxu.org/) (Stanford); [Eddie Yang](https://www.eddieyang.net/) (UCSD)

**Date:** March 17, 2022

**Version:** 1.2.2 ([Github](https://github.com/xuyiqing/hbal)); coming soon ([CRAN])

**Reference:** Xu, Yiqing and Eddie Yang (2021). "Hierarchically Regularized Entropy Balancing" Available at [SSRN](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3807620).

R code used in this demonstration can be downloaded from [here](http://yiqingxu.org/software/xuyiqing/hbal_examples.R).

***
**Updates in v.1.2.2**

First CRAN release!

***

## Contents

1. Installation

2. Balancing Covariates

3. Visualizing Results

4. More Options

5. Relation to **ebal**

***

## 1. Installation

You can install the **hbal** package from CRAN: 
```{r eval=FALSE}
install.packages('hbal') # Not run. Not yet on CRAN.
```

You can also install the up-to-date development version from Github: 
```{r eval=FALSE}
install.packages('devtools', repos = 'http://cran.us.r-project.org') # if not already installed
devtools::install_github('xuyiqing/hbal', ref="main")
```

<!--
**hbal** depends on the following packages, which will be installed AUTOMATICALLY when **hbal** is being installed; you can also install them manually:  
```{r eval=FALSE}
require(estimatr)  
require(glmnet) 
require(ggplot2)
require(gridExtra)
require(gtable)
require(nloptr)
require(Rcpp)
require(RcppEigen)
require(stringr)
```-->

***
## 2. Balancing Covariates

We simulate a toy cross-sectional dataset with binary treatment before showing how to use **hbal**. 
```{r, message=FALSE}
library(hbal)
set.seed(92092)
N <- 500
X1 <- rnorm(N)
X2 <- rbinom(N,size=1,prob=.5)
X <- cbind(X1, X2) # Covariates
treat <- rbinom(N, 1, prob=0.5) # Treatment indicator
y <- X[,1] + X[,2] + rnorm(N) # Outcome
```

In the simplest use case, we can use the following two lines of code to:

1. Bbalance the covariates (up to the third moments) between the treatmeant and control groups.

2. estimate the Average Treatment Effect on the Treated (ATT).

```{r}
out <- hbal(Treatment = treat, X = X,  Y = y)
summary(hbal::att(out))
```

We can see that, by default, `hbal` balances on an expanded covariate set that includes second- and third-order polynomials of the covariates. 

`att` uses linear regression with robust standard errors (`lm_robust` from the **estimatr** package) to calculate the ATT. Additional arguments accepted by `lm_robust`, such as clusters, can be passed to `att()`.

`hbal` returns a list of 9 objects:

```{r}
objects(out)
```
1. **call**: A string of the function call
2. **coefs**: Values of Lagrangian multipliers. They are used to calculate the solution `weights`.
3. **converged**: Binary indicator of whether the algorithm has converged.
4. **group.assignment**: A vector of the number of variabels in each covariate group.
5. **mat**: Expanded covariates matrix.
6. **penalty**: This is the regularization parameter $\alpha$ in  Xu \& Yang (2021).
7. **Treatment**: Treatment indicator. Reproduced here to be used by `att()`.
8. **weights**: Solution weights. Can be plugged into any downstream estimator.
9. **Y**: Outcome variable. Reproduced here to be used by `att()`.


***
## 3. Visualizing Results

`hbal` has a build-in `plot` method that allows us to visualize covariate balance before and after balancing.

```{r, fig.height = 5.5, fig.width = 9, fig.align = "left"}
plot(out)
```

We can see that the linear terms (mean) of the covariates are exactly balanced between the treatment and the control groups. We can check the penalties applied to different groups of covariates. In this case, the higher-order terms have relatively high penalties, meaning that `hbal` determines that they are not informative of the treatment assignment process (which is true of the actual treatment assignment) and some imbalance can be tolerated. 

```{r}
out$penalty
```

We can also plot the weight distribution for the control units by specifying `type = 'weight'` in `plot`. We can see that the weights are quite evenly distributed around the unit weight.

```{r}
plot(out, type='weight')
```

***
## 4. More Options

### 4.1 Setting Series Expansion

`hbal` uses the R built-in function `poly` to include higher-order polynomials of the supplied covariates in the balancing scheme. This is controlled by the `expand.degree` argument. By default, it is set to `expand.degree = 3`, which expands the covariates to include polynomials up to the 3rd degree.

We can ask `hbal` to balance on less flexible functions of the covariates by decreasing the value of `expand.degree`, e.g. by setting `expand.degree = 2`. This may be useful when balancing on third order terms is infeasible.

***

### 4.2 Excluding Covariates

By default, `hbal` uses the R built-in `qr` to check the rank of the (expanded) covariate matrix and remove columns that are not pivots when the matrix is rank-deficient. However, if a priori we know some combinations of the covariates are nonsensical, we can exclude them explicitly by using the `exclude` argument. For example, we can exclude any interaction that involves **X1** and **X2**:

```{r}
out <- hbal(Treatment = treat, Y = y, X = X, exclude=list(c("X1", "X2")))
summary(att(out)) # X1.X2 and X1.X1.X2 removed from balancing scheme
```

***

### 4.3 Custom K-fold cross-validation

By default, `hbal` uses 4-fold cross-validation. We can change to K-fold cross-validation for any arbitrary K by setting `folds = K`.

We can also disable cross-validation by setting `cv = FALSE`. No regularization will be applied in this case and `hbal` is essentially equivalent to `ebalance` from the **ebal** package.

***

### 4.4 User-supplied base weights

By default, `hbal` tries to keep the solution weights for the control units as close as possible (in an entropy sense) to a set of uniform base weights to retain information. In cases where the target distribution of the weights for the control units is not uniform weights, we can incorporate this information by supplying a vector of target weights to `base.weight`. 

For example, if we want to set the target weight distribution such that the first 100 control units have weights of $\frac{1}{2}$ while the rest of the control units have weights of 1, we can do:

***

### 4.5 Other functionalities

1. `ds`: The double selection method by Belloni, Chernozhukov and Hansen (2014). This screens the expanded covariates and only keeps those that are are predictive for the treatment assignment or the outcome. This further reduces the dimensionality of the problem. Default is set to `FALSE`.

2. `constraint.tolerance`: Convergence criterion. The optimization algorithm will stop when the maximum difference in covaraite means between the treated and the control units is below `constraint.tolerance`.

3. `shuffle.treat`: Whether treated units should be partitioned in cross-validation. Default is set to `FALSE`. If set to `TRUE`, the covariate means of the treated units will vary from fold to fold. It may be advisable to set this argument to `FALSE` if there is only a small number of treated units or if there are many outliers in the treatment group.

4. `max.iterations`: Maximum number of iterations that will be run in finding the solution weights. Default is set to 200.

## 5. Relation to **ebal**
By setting `expand.degree=0` and `cv=FALSE`, which tells `hbal` to not expand the covariate space and not use cross validation to search for regularization hyperparameters, `hbal` is equivalent to entropy balancing (Hainmueller, 2012). We can demonstrate this equivalence by showing the hbal weights are exactly the same to the ebal weights from the **ebal** package in this case, as seen in the plot below.

```{r}
library(ebal)

ebal_mean_balance <- ebalance(Treatment=treat, X=X) # mean balancing only
hbal_mean_balance <- hbal(Treatment=treat, X=X, Y=y, expand.degree=0, cv=FALSE) # mean balancing only
W <- data.frame(x=hbal_mean_balance$weights*sum(treat), 
                y=ebal_mean_balance$w) # store weights as x-y coordinates
ggplot(aes(x=x, y=y), data=W) + geom_point() + theme_bw() + 
  labs(x="hbal weights", y="ebal weights", title="correlation between ebal and hbal weights")
```

***
## Reference

Belloni, A., Chernozhukov, V., & Hansen, C. (2014). Inference on treatment effects after selection among high-dimensional controls. The Review of Economic Studies, 81(2), 608-650.

Black, R. C., & Owens, R. J. (2016). Courting the president: how circuit court judges alter their behavior for promotion to the Supreme Court. American Journal of Political Science, 60(1), 30-43.

Hainmueller, J. (2012). Entropy balancing for causal effects: A multivariate reweighting method to produce balanced samples in observational studies. Political analysis, 25-46.

Lin, W. (2013). Agnostic notes on regression adjustments to experimental data: Reexamining Freedman’s critique. Annals of Applied Statistics, 7(1), 295-318.

***
Please report bugs and let us know if you have any suggestions! -> z5yang [at] ucsd.edu
