---
title: 'Tutorial'
output:
  rmarkdown::html_vignette:
    toc: false
    toc_depth: 4
    number_sections: false
bibliography: references.bib      
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---
<!-- 
  Code to Justify Text
    <style>
    body {
    text-align: justify}
    </style>
-->   
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 10,
  fig.height = 7,
  comment = "#>"
)
``` 


```{r, include = FALSE}
library(ggplot2)
library(estimatr)
require(ebal)
```

This page demonstrates the usage of the **hbal** package, which implements hierarchically regularized entropy balancing introduced by @XuYang2022.  **hbal** automatically expands the covariate space to include higher order terms and uses cross-validation to select variable penalties for the balancing conditions, and then seeks approximate balance on the expanded covariate space. 

**hbal** provides two main functions:

* `hbal()`: performs hierarchically regularized entropy balancing.

* `att()`: calculates the average treatment effect on the treated (ATT) from an `hbal` object returned by `hbal()`. 

And two S3 methods:

* `summary()`: summarizes the balancing results. 

* `plot()`: visualizes covariate balance before and after weighting or the distribution of balancing weights.

R code used in this tutorial can be downloaded from [here](https://raw.githubusercontent.com/xuyiqing/hbal/main/pkgdown/hbal_examples.R).

***

## Basic Usage

We simulate a toy cross-sectional dataset with a binary treatment to illustrate the basic usage of **hbal**. Note that treatment assignment depends on all three covariates. 
```{r, message=FALSE}
library(hbal)
set.seed(1984)
N <- 1500
X1 <- rnorm(N)
X2 <- rnorm(N)
X3 <- rbinom(N, size = 1, prob = .5)
D_star <- 0.5 * X1 + 0.3 * X2 + 0.2 * X1 * X2 - 0.5 * X1 * X3 - 1
D <- ifelse(D_star > rnorm(N), 1, 0) # Treatment indicator
y <- 0.5 * D + X1 + X2 + X2 * X3 + rnorm(N) # Outcome
dat <- data.frame(D = D, X1 = X1, X2 = X2, X3 = X3, Y = y)
head(dat)
```

**hbal** is an extension of entropy balancing, or **ebal** proposed by @hainmueller2012entropy. By default, **hbal** replicates **ebal** by performing exact balancing on all covariates and no serial expansion, i.e., `expand.degree = 1` (default). We can demonstrate this equivalence by showing the hbal weights are exactly the same to the ebal weights from the **ebal** package in this case. 

```{r, fig.height = 6, fig.width = 6, fig.align = "left", dpi=100}
library(ebal)
ebal.out <- ebalance(Treat = dat$D, X = dat[,c('X1', 'X2', 'X3')]) # ebal
hbal.out <- hbal(Treat = 'D', X = c('X1', 'X2', 'X3'),  Y = 'Y',  data = dat) # hbal

# plot weights
W <- data.frame(x = ebal.out$w, y = hbal.out$weights.co)
ggplot(aes(x = x, y = y), data = W) + geom_point() + theme_bw() + 
  labs(x = "ebal weights", y="hbal weights", title = "ebal weights vs. hbal weights")
```

`hbal()` returns a list of 15 objects:

```{r}
names(hbal.out)
```
1. **converged**: Binary indicator of whether the algorithm has converged.
2. **weights**: Resulting weights, including the base weights for the treated and solution weights for the controls (see below).
3. **weights.co**: Solution weights for the controls. Can be plugged into any downstream estimator.
4. **coefs**: Values of Lagrangian multipliers. They are used to calculate the solution `weights`.
5. **Treatment**: The treatment status vector.
6. **mat**: Data matrix, including the outcome variable (the first column), the treatment variable (the second column), and expanded covariates (the remaining columns).
7. **grouping**: A vector of the number of variables in each covariate group. The length of the vector equals the number of groups.
8. **group.penalty**: Penalties for different groups of covariates. This is the regularization parameter $\alpha$ in @XuYang2022.
9. **term.penalty**: Penalties for individual covariates. The length of the vector equals the number of balancing terms, including expanded ones.
10. **bal.tal**: A balance table.
11. **base.weights**: A vector of base weights, originally from variable `w` supplied by users. The length of the vector equals the number of rows entering the balancing scheme.
12. **Treat**: A character string that stores the treatment variable name. 
13. **Outcome**: The outcome vector, if `Y` is supplied.
14. **Y**: A character string that stores the outcome variable name, if `Y` is supplied.. 
15. **call**: A string of the function call.

The `summary()` function provides additional information on the balancing scheme, including the numbers of treated and control units, groups and corresponding penalties, and a balance table, in which `Std.Diff.(O)` and `Std.Diff.(W)` represent standardized difference before and after balancing, respectively. For example, in this case, only the linear terms are being balanced on. 
```{r}
summary(hbal.out)
```

In addition to balancing on just the linear/level terms of the covariates, `hbal()` allows balancing on a serial expansion of the covariates. This is achieved by setting the `expand.degree` argument. Currently, `hbal()` supports both second order serial expansion (two-way interactions and square terms; `expand.degree = 2`) and third order serial expansion (two-way interactions, square terms, linear*square interactions, and cubic terms; `expand.degree = 3`). For example, we can do exact balancing on third order serial expansion of the covariates:
```{r}
out <- hbal(Y = 'Y', Treat = 'D', X = c('X1', 'X2', 'X3'),  
            data = dat, expand.degree = 3)
summary(out)
```

### Obtaining the ATT

We can use `att()` on a `hbal` object to directly get an ATT estimate.  Note that `att()` uses linear regression with robust standard errors (`lm_robust()` from the **estimatr** package) to calculate the ATT. `att()` allows users to use one of the following three methods to estimate the ATT:

* difference in weighted means by setting `dr = FASLE`, in which `dr` represents *doubly robust*; as a result, `method` is ignored;

* weighted linear regression by setting `method = ""lm_robust"` (default);

* weighted linear regression with a full set of interactions between the treatment indicator and demeaned covariates, or Lin's method, by setting `method = ""lm_lin"`.  

Additional arguments accepted by `lm_robust()` or `lm_lin()`, such as `se_type` and `clusters`, can be passed to `att()`, for example: 
```{r}
att(out, dr = FALSE)
att(out)
att(out, method = "lm_lin", se_type = "stata")
```


Furthermore, `hbal()` allows variable penalties on the balancing conditions for different groups of covariates. Penalties are automatically determined using cross-validation. This minimizes the variance of the weights, increases the feasibility of the balancing problem, and is the approach advocated in @XuYang2022. We can do so by setting `cv = TRUE`. Note that, by default, `hbal()` only penalizes higher-order terms and seeks to achieve exact balance on linear terms.

```{r}
out <- hbal(Treat = 'D', X = c('X1', 'X2', 'X3'),  Y = 'Y', 
            data = dat, expand.degree = 3, cv = TRUE)
summary(out)
att(out)
```


### Visualizing Results

**hbal** has a build-in `plot()` method that allows us to visualize covariate balance before and after balancing.

```{r, fig.align = "left"}
plot(out)
```

From the above plots, we can see that the linear terms of the covariates are exactly balanced between the treatment and the control groups. We can check the group penalties or penalties applied to each covariate. In this case, the higher-order terms all have relatively high penalties except for two-way interactions, in accordance with the true data generating process. 

```{r}
round(out$group.penalty, 2)
round(out$term.penalty, 2)
```

We can also plot the weight distribution for the control units by specifying `type = 'weight'` in `plot()`. We can see that the weights are quite concentrated around the unit weight.

```{r, fig.width = 6, fig.height = 6, fig.align = "left"}
plot(out, type='weight')
```


## Additional Options
### Controlling Exact/Approximate Balancing

Users can manually control which group of covariates to be exactly balanced and which group to be approximately balanced (via regularization) by using the `group.exact` argument. It takes in a binary vector with length equal to the number of covariate groups, with 1 indicating exact balance on that group and 0 indicating approximate balance. Appropriate penalties are then searched through cross-validation. When using this option, please make that the length of vector supplied equals the number of groups. 

For example, we can ask `hbal()` to achieve exact balance on the linear terms and two-way interactions and approximate balance for the rest:
```{r}
out <- hbal(Treat = 'D', X = c('X1', 'X2'),  Y = 'Y', data = dat, 
            expand.degree = 3, cv = TRUE, group.exact = c(1, 1, 0, 0, 0))
summary(out)
```

### User-Supplied Penalties
If researchers have prior knowledge about covariates and would like to supply specific penalties for each group of covariates, we can do so by passing a vector of penalties to the `group.alpha` argument. For example, we can manually set penalties to be 0 for the the linear terms and two-way interactions and 100 for other groups:
```{r}
out <- hbal(Treat = 'D', X = c('X1', 'X2', 'X3'),  Y = 'Y', data = dat, 
            expand.degree = 3, group.alpha = c(0, 0, 100, 100, 100, 100))
summary(out)
```

### Controling Serial Expansion

Sometimes we may not want to perform serial expansion on all covariates but instead expand on a select set of covariates, we can do so by using the `X.expand` argument to specify the covariates we want to expand on. For example, we can expand only on **X1**:
```{r}
out <- hbal(Treat = 'D', X = c('X1', 'X2', 'X3'),  Y = 'Y', data = dat, 
            expand.degree = 3, X.expand = c('X1', 'X2'))
summary(out)
```

### Selecting Covariates
Performing serial expansion on many covariates can result in a prohibitive number of covariates that need to be balanced on. In such cases, users many want to reduce the number of covariate by using the double selection method by @belloni2014inference. This screens the expanded covariates and only keeps those that are are predictive for the treatment assignment or the outcome. Users can enable double selection by setting `ds = TRUE`. In the following case, almost all the higher-order terms except `X1*X2` and `X2*X3` are screened out by the double selection procedure.
```{r}
out <- hbal(Treat = 'D', X = c('X1', 'X2', 'X3'),  Y = 'Y', data = dat, 
            expand.degree = 3, ds = TRUE) 
summary(out)
att(out)
```

### Keeping/Excluding Covariates
If there are covariates that users would like to keep in the balancing conditions regardless of whether they are selected in double selection, we can use the `X.keep` argument to specify the covariates to keep. On the other hand, if a priori we know some combinations of the covariates are nonsensical, we can exclude them explicitly by using the `exclude` argument. For example, we can exclude any interaction that involves **X1** and **X2**:
```{r}
out <- hbal(Treat = 'D', X = c('X1', 'X2', 'X3'),  Y = 'Y', data = dat, 
            expand.degree = 3, exclude = list(c('X1', 'X2')))
summary(out) # X1.X2 and X1.X1.X2 removed from balancing scheme
att(out) 
```

## Example 1: Lalonde Data 

To show the use of **habl** in practice, here we use **habl** on a dataset that contains the subset of the @lalonde1986evaluating dataset from @dehejia1999causal and the Panel Study of Income Dynamics (PSID-1), which is also shipped with **hbal**. 
```{r}
data(hbal)
head(lalonde)
```

First, we adjust for the linear terms only:
```{r}
xvars <- c("age","black","educ","hisp","married","re74","re75","nodegr","u74","u75") # covariates
# hbal w/ level terms only
hbal.out <- hbal(Treat = 'nsw', X = xvars,  Y = 're78', data = lalonde) 
summary(hbal.out)
att(hbal.out)
```

Adding higher-order terms makes the treatment effect estimate closer to the experimental benchmark (~$1800). As a result of cross-validation, no penalty is imposed on the linear and squared terms.  Note that `exclude=list(c("educ", "nodegr"))` removes the nonsensical interaction between `educ` and `nodegr`.
```{r}
hbal.full.out <- hbal(Treat = 'nsw', X = xvars, Y = 're78', data = lalonde, 
                      expand.degree = 2, cv = TRUE, exclude=list(c("educ", "nodegr")))
summary(hbal.full.out)
att(hbal.full.out)
```

We can check the penalties applied to each group and the cavariate balance before and after balancing.
```{r, fig.height = 8, fig.align = "left"}
hbal.full.out$group.penalty
plot(hbal.full.out)
```

## Example 2: Black and Owens (2016)

The second example comes from @black2016courting, in which the authors study the effect of promotion prospect to the Supreme Court on the behavior of circuit court judges. Here we focus on whether circuit court judges who are on the shortlist to fill Supreme Court vacancies ("contenders") ruled in line with the president as the outcome of interest. We load the dataset `contenderJudges`, which is shipped with **hbal**.
```{r}
data(hbal)
```

First, we take a look at the data structure. The outcome variable is `presIdeoVote` and the treatment variable is `treatFinal0`, indicating whether there was a Supreme Court vacancy at the time. There are also 7 covariates on judge and court characteristics and a variable `judge` that indicates the judges' names.
```{r}
str(contenderJudges)
```

We can estimate the effect of Supreme Court vacancy on judges' rulings while controlling for functions of the covariates (to the second order). We see that contender judges are more likely to rule in line with the ideology of the sitting president during a Supreme Court vacancy.

```{r}
xvars <- c("judgeJCS", "presDist", "panelDistJCS", "circmed", "sctmed", "coarevtc", "casepub")
out <- hbal(Treat = 'treatFinal0', X = xvars, Y = 'presIdeoVote', data = contenderJudges,
            expand.degree = 2, cv = TRUE)
summary(out)
att(out)
```

We can further check covariate balance before and after balancing by checking the balance plots. Here we see that the linear terms are exactly balanced between the treatment and the control groups. Imbalance among higher-order terms and interactions are also significantly reduced.

```{r, fig.height = 8, fig.align = "left"}
plot(out)
```

# Reference


