---
title: 'Tutorial'
output:
  rmarkdown::html_vignette:
    toc: false
    toc_depth: 4
    number_sections: false
bibliography: references.bib      
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---
<!-- 
  Code to Justify Text
    <style>
    body {
    text-align: justify}
    </style>
-->   
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 10,
  fig.height = 7,
  comment = "#>"
)
``` 


```{r, include = FALSE}
library(ggplot2)
library(estimatr)
require(ebal)
```

This page demonstrates the usage of the **hbal** package, which implements hierarchically regularized entropy balancing introduced by @XuYang2022.  **hbal** automatically expands the covariate space to include higher order terms and uses cross-validation to select variable penalties for the balancing conditions, and then seeks approximate balance on the expanded covariate space. 

**hbal** provides two main functions:

* `hbal()`: performs hierarchically regularized entropy balancing.

* `att()`: calculates the average treatment effect on the treated (ATT) from an `hbal` object returned by `hbal()`. 

And two S3 methods:

* `summary()`: summarizes the balancing results. 

* `plot()`: visualizes covariate balance before and after weighting or the distribution of balancing weights.

R code used in this tutorial can be downloaded from [here](https://raw.githubusercontent.com/xuyiqing/hbal/main/pkgdown/hbal_examples.R).

***

## Basic Usage

We simulate a toy cross-sectional dataset with a binary treatment to illustrate the basic usage of **hbal**. Note that treatment assignment depends on all three covariates. 
```{r, message=FALSE}
library(hbal)
set.seed(1984)
N <- 1500
X1 <- rnorm(N)
X2 <- rnorm(N)
X3 <- rbinom(N, size = 1, prob = .5)
D_star <- 0.5 * X1 + 0.3 * X2 + 0.2 * X1 * X2 - 0.5 * X1 * X3 - 1
D <- ifelse(D_star > rnorm(N), 1, 0) # Treatment indicator
y <- 0.5 * D + X1 + X2 + X2 * X3 + rnorm(N) # Outcome
dat <- data.frame(D = D, X1 = X1, X2 = X2, X3 = X3, Y = y)
head(dat)
```

**hbal** is an extension of entropy balancing, or **ebal** proposed by @hainmueller2012entropy. By default, **hbal** replicates **ebal** by performing exact balancing on all covariates and no serial expansion, i.e., `expand.degree = 1` (default). We can demonstrate this equivalence by showing the hbal weights are exactly the same to the ebal weights from the **ebal** package in this case. 

```{r, fig.height = 6, fig.width = 6, fig.align = "left", dpi=100}
library(ebal)
ebal.out <- ebalance(Treat = dat$D, X = dat[,c('X1', 'X2', 'X3')]) # ebal
hbal.out <- hbal(Treat = 'D', X = c('X1', 'X2', 'X3'),  Y = 'Y',  data = dat) # hbal

# plot weights
W <- data.frame(x = ebal.out$w, y = hbal.out$weights.co)
ggplot(aes(x = x, y = y), data = W) + geom_point() + theme_bw() + 
  labs(x = "ebal weights", y="hbal weights", title = "ebal weights vs. hbal weights")
```

`hbal()` returns a list of 15 objects:

```{r}
names(hbal.out)
```
1. **converged**: Binary indicator of whether the algorithm has converged.
2. **weights**: Resulting weights, including the base weights for the treated and solution weights for the controls (see below).
3. **weights.co**: Solution weights for the controls. Can be plugged into any downstream estimator.
4. **coefs**: Values of Lagrangian multipliers. They are used to calculate the solution `weights`.
5. **Treatment**: The treatment status vector.
6. **mat**: Data matrix, including the outcome variable (the first column), the treatment variable (the second column), and expanded covariates (the remaining columns).
7. **grouping**: A vector of the number of variables in each covariate group. The length of the vector equals the number of groups.
8. **group.penalty**: Penalties for different groups of covariates. This is the regularization parameter $\alpha$ in @XuYang2022.
9. **term.penalty**: Penalties for individual covariates. The length of the vector equals the number of balancing terms, including expanded ones.
10. **bal.tal**: A balance table.
11. **base.weights**: A vector of base weights, originally from variable `w` supplied by users. The length of the vector equals the number of rows entering the balancing scheme.
12. **Treat**: A character string that stores the treatment variable name. 
13. **Outcome**: The outcome vector, if `Y` is supplied.
14. **Y**: A character string that stores the outcome variable name, if `Y` is supplied.. 
15. **call**: A string of the function call.

The `summary()` function provides additional information on the balancing scheme, including the numbers of treated and control units, groups and corresponding penalties, and a balance table, in which `Std.Diff.(O)` and `Std.Diff.(W)` represent standardized difference before and after balancing, respectively. For example, in this case, only the linear terms are being balanced on. 
```{r}
summary(hbal.out)
```

In addition to balancing on just the linear/level terms of the covariates, `hbal()` allows balancing on a serial expansion of the covariates. This is achieved by setting the `expand.degree` argument. Currently, `hbal()` supports both second order serial expansion (two-way interactions and square terms; `expand.degree = 2`) and third order serial expansion (two-way interactions, square terms, linear*square interactions, and cubic terms; `expand.degree = 3`). For example, we can do exact balancing on third order serial expansion of the covariates:
```{r}
out <- hbal(Y = 'Y', Treat = 'D', X = c('X1', 'X2', 'X3'),  
            data = dat, expand.degree = 3)
summary(out)
```

### Obtaining the ATT

We can use `att()` on an `hbal` object to get an ATT estimate.

#### The default estimator: augmented balancing weights (`method = "abw"`)

Since version 1.3.0, the default estimator is an augmented balancing weights estimator (`method = "abw"`; @benmichael2021balancing, @brunssmith2023augmented). It combines the hbal weights with an outcome model in three steps:

1. Fit a regression of the outcome on the columns of `mat`, using the control units only and weighted by their hbal weights. The regression is a ridge regression on standardized columns with an unpenalized intercept. Its penalty is chosen once from the full control sample by generalized cross-validation [@golub1979generalized] over a fixed grid, with a cap on the effective degrees of freedom so that the fit can never interpolate the controls. No random numbers are involved.

2. Predict the untreated outcome of each treated unit from that regression. Then correct the prediction with the hbal-weighted average of the control residuals. The residual term corrects the regression where it is wrong; the regression corrects the weights where the balance is only approximate.

3. Take the weighted mean of the treated outcomes and subtract the corrected prediction.

The resulting estimator is doubly robust. It is also Neyman-orthogonal: a small error in either the outcome regression or the weights has no first-order effect on the estimate. The control residuals in step 2 are cross-fitted: the controls are split into folds, and each control's residual comes from a regression fitted without its own fold. Standard errors come from the influence function of the estimator, treating the weights and the outcome regression as fixed; the p-value and the 95 percent confidence interval use a t distribution with n − 1 degrees of freedom.

If the entropy-balancing algorithm did not converge (`out$converged` is 0), `att()`'s default method warns that the estimate and its standard error may be unreliable. The warning names the usual remedies: refit `hbal()` with `cv = TRUE`, a lower `expand.degree`, or the `ds`, `X.keep`, or `exclude` options.

#### Arguments

| Argument | Default | What it does |
| --- | --- | --- |
| `method` | `"abw"` | Which estimator to use: `"abw"` (default, above), `"lm_robust"`, `"lm_lin"`, or `"elnet"` (below). |
| `dr` | `TRUE` | Whether to include the outcome model. `FALSE` drops it: `"abw"` and `"lm_robust"` then both return the weighted difference in means. |
| `nfolds` | `5` | Number of cross-fitting folds for `method = "abw"`. `att()` uses fewer, and says so in a message, when the control group is small relative to the number of columns in `mat`. Ignored by the other methods. |
| `seed` | `NULL` | Fixes the cross-fitting fold assignment for `method = "abw"`. The assignment is a deterministic function of `seed`; `att()` never calls `set.seed()`, so results are identical across repeated calls with or without a seed. Ignored by the other methods. |
| `displayAll` | `FALSE` | `FALSE` returns the one-row summary table. `TRUE` returns the full result instead: a list for `"abw"` (see above), the fitted model object for `"lm_robust"`/`"lm_lin"`. |
| `...` | — | Extra arguments, forwarded to `lm_robust()`/`lm_lin()` (e.g. `se_type`, `clusters`). `method = "abw"` does not accept any and stops with an error naming them. |

#### The legacy estimators

The three estimators available before version 1.3.0 remain available through the `method` argument and give exactly the numbers they gave before:

* `method = "lm_robust"`: weighted linear regression with robust standard errors, using `lm_robust()` from the **estimatr** package. This was the default before version 1.3.0. Set it explicitly to reproduce results from earlier versions.

* `method = "lm_lin"`: weighted linear regression with a full set of interactions between the treatment indicator and the demeaned covariates, or Lin's method [@lin2013agnostic], using `lm_lin()` from **estimatr**.

* `method = "elnet"`: approximate residual balancing [@athey2018approximate], which fits an elastic-net outcome model with **glmnet**.

Additional arguments accepted by `lm_robust()` or `lm_lin()`, such as `se_type` and `clusters`, can be passed to `att()` with those two methods only. The default method does not accept them and stops with an error that points to `method = "lm_robust"` or `"lm_lin"`. Setting `dr = FALSE` drops the outcome model. With `dr = FALSE`, the default method and `method = "lm_robust"` both return the difference in weighted means; they differ only in how the standard error is computed. Note that `att(out, dr = FALSE)` now uses the default method, so add `method = "lm_robust"` to reproduce the standard errors from earlier versions.

```{r}
att(out)
att(out, method = "lm_robust")
att(out, method = "lm_lin", se_type = "stata")
att(out, dr = FALSE)
```

By default, `att()` only displays the treatment effect (`displayAll = FALSE`). With `displayAll = TRUE`, the default method returns a list with the estimate, its standard error, the per-unit influence values, the fold assignment, the coefficients of the outcome regression, and the weights. For `method = "lm_robust"` or `"lm_lin"`, `displayAll = TRUE` returns the fitted **estimatr** object instead.
```{r}
res <- att(out, displayAll = TRUE)
str(res, max.level = 1)
round(res$nuisance$coef_full, 3)
```

#### Which method should I use?

* `method = "abw"` (the default) for inference: its 95 percent coverage is 0.945 to 0.956 across the sample sizes and the two main scenarios in the simulations below.

* `method = "lm_robust"`, with no other arguments changed, to reproduce a result computed with hbal before version 1.3.0 exactly — point estimate and standard error both.

* `method = "lm_robust", se_type = "HC3"` for a new small-sample analysis (N of 500 or fewer) that uses the pre-1.3.0 point estimate but wants a better-calibrated standard error than its pre-1.3.0 default provides; see below for why.

* `nfolds = 1` only when a smaller root-mean-squared error matters more than a reliable standard error, and the sample is small; see below for the trade-off.

* `method = "elnet"` for a high-dimensional expansion (many covariates, or a high `expand.degree`); this method was not part of the simulations below.

#### Evidence from simulations

Simulations accompanying the 1.3.0 release compare `method = "abw"` (the default) against `method = "lm_robust"` (the pre-1.3.0 default) across three data-generating processes and three sample sizes (N = 200, 500, 1000), 1,000 replications per cell.

**Small samples.** At N = 200, the default's root-mean-squared error (RMSE) is higher than `lm_robust`'s: 0.216 vs 0.189 (14 percent higher) when the outcome model is right and the weights are only approximately balanced, and 0.227 vs 0.169 (34 percent higher) when both the weights and the outcome model are mildly wrong. The gap narrows as N grows: by N = 1000 it is 0.087 vs 0.086 and 0.082 vs 0.080, both under 3 percent. In exchange, the default's standard errors are calibrated when the outcome model is right: an SE ratio (mean estimated standard error divided by the actual spread of the estimates across replications) of 0.99 to 1.00 and 95 percent coverage of about 0.95 (0.948 to 0.956), at every N. `lm_robust`'s standard error there runs 5 to 7 percent short at every N, not only at N = 200 (SE ratio 0.93 to 0.95; coverage 0.93 to 0.945).

**`nfolds = 1` (no cross-fitting).** Setting `nfolds = 1` turns off cross-fitting, and the residual-correction term in step 2 becomes identically zero: what `att()` returns is then exactly the hbal-weighted ridge regression's prediction of the treated units' counterfactual outcome, with no correction step. At N = 200 this closes most of the small-sample RMSE gap to `lm_robust`: RMSE falls from 0.216 to 0.200 when the outcome model is right (59 percent of the gap to `lm_robust`'s 0.189), and from 0.227 to 0.190 when both nuisances are mildly wrong (64 percent of the gap to `lm_robust`'s 0.169). The cost is the standard error: at N = 200 the SE ratio drops to 0.93 (from the default's 0.995) when the outcome model is right, and to 0.89 (from 0.924) when both nuisances are wrong, and 95 percent coverage falls to 0.935 and 0.921 respectively (both below the default's 0.956). Under the mildly misspecified scenario, `nfolds = 1`'s bias is also larger than the default's at every N (0.031 to 0.033 vs 0.011 to 0.025), because cross-fitting is what supplies the finite-sample correction that shrinks it. Ten folds do no better than the default five — RMSE stays within 2.3 percent of the five-fold value at every N in these two scenarios — so five remains the default.

**Standard errors for `method = "lm_robust"`.** Its default standard error (`se_type = "stata"`, equivalent to HC1 with no clusters) runs short when the outcome model is right: an SE ratio of 0.93 to 0.95, 4.7 to 7.2 percent short of calibrated, at every N tested. Setting `se_type = "HC3"` closes most of this gap at N of 500 or fewer (SE ratio 0.997 to 1.042, coverage 0.949 to 0.956) but still leaves the standard error about 5 percent short at N = 1000 (SE ratio 0.951, coverage 0.938). This remaining shortfall is not a small-sample leverage effect: the leverage correction itself shrinks with N as it should (HC3 adds only 2.5 percent to the standard error at N = 1000, against 9.7 percent at N = 200), while the part of the shortfall that survives at N = 1000 does not shrink that way. A plausible source, not tested here, is that the standard-error formula treats the estimated hbal weights as fixed.

The table below compares the default and `lm_robust` directly, at N = 200 and N = 1000, in the two scenarios above (true ATT: 2 in the first, 1.593 in the second; N = 500, not shown, gives the same coverage ordering: 0.954 vs 0.941 in the first scenario, 0.947 vs 0.935 in the second).

| Scenario | N | Method | Bias | SE ratio | 95% coverage |
| --- | --- | --- | --- | --- | --- |
| Weights approximate, outcome model right | 200 | abw (default) | -0.018 | 0.995 | 0.956 |
| Weights approximate, outcome model right | 200 | lm_robust | 0.001 | 0.950 | 0.945 |
| Weights approximate, outcome model right | 1000 | abw (default) | -0.001 | 0.992 | 0.948 |
| Weights approximate, outcome model right | 1000 | lm_robust | 0.000 | 0.928 | 0.929 |
| Both nuisances mildly wrong | 200 | abw (default) | 0.011 | 0.924 | 0.956 |
| Both nuisances mildly wrong | 200 | lm_robust | 0.023 | 0.984 | 0.941 |
| Both nuisances mildly wrong | 1000 | abw (default) | 0.025 | 1.012 | 0.945 |
| Both nuisances mildly wrong | 1000 | lm_robust | 0.030 | 0.989 | 0.927 |

Both methods are close to unbiased when only the weights are approximately balanced; when the outcome model is also wrong, the default's bias is about half of `lm_robust`'s at N = 200 (0.011 vs 0.023) but its standard error is the less calibrated of the two there (SE ratio 0.924 vs 0.984). By N = 1000 the default's standard error is calibrated in both scenarios (0.99 to 1.01) while `lm_robust`'s remains 1 to 7 percent short and its coverage 0.927 to 0.929, short of the nominal 95 percent.

A third scenario (not shown above) makes the confounding a nonlinear function of the observed covariates, severe enough that no polynomial expansion of `mat` can represent it. There, neither estimator is consistent, but the default's bias is smaller at every N (1.74 vs 2.32 at N = 200) while its spread is larger (SD 8.89 vs 2.46 at N = 200, narrowing to 2.28 vs 1.01 by N = 1000).

Furthermore, `hbal()` allows variable penalties on the balancing conditions for different groups of covariates. Penalties are automatically determined using cross-validation. This minimizes the variance of the weights, increases the feasibility of the balancing problem, and is the approach advocated in @XuYang2022. We can do so by setting `cv = TRUE`. Note that, by default, `hbal()` only penalizes higher-order terms and seeks to achieve exact balance on linear terms.

```{r}
out <- hbal(Treat = 'D', X = c('X1', 'X2', 'X3'),  Y = 'Y', 
            data = dat, expand.degree = 3, cv = TRUE)
summary(out)
att(out)
```


### Visualizing Results

**hbal** has a build-in `plot()` method that allows us to visualize covariate balance before and after balancing.

```{r, fig.align = "left"}
plot(out)
```

From the above plots, we can see that the linear terms of the covariates are exactly balanced between the treatment and the control groups. We can check the group penalties or penalties applied to each covariate. In this case, the higher-order terms all have relatively high penalties except for two-way interactions, in accordance with the true data generating process. 

```{r}
round(out$group.penalty, 2)
round(out$term.penalty, 2)
```

We can also plot the weight distribution for the control units by specifying `type = 'weight'` in `plot()`. We can see that the weights are quite concentrated around the unit weight.

```{r, fig.width = 6, fig.height = 6, fig.align = "left"}
plot(out, type='weight')
```


## Additional Options
### Controlling Exact/Approximate Balancing

Users can manually control which group of covariates to be exactly balanced and which group to be approximately balanced (via regularization) by using the `group.exact` argument. It takes in a binary vector with length equal to the number of covariate groups, with 1 indicating exact balance on that group and 0 indicating approximate balance. Appropriate penalties are then searched through cross-validation. When using this option, please make that the length of vector supplied equals the number of groups. 

For example, we can ask `hbal()` to achieve exact balance on the linear terms and two-way interactions and approximate balance for the rest:
```{r}
out <- hbal(Treat = 'D', X = c('X1', 'X2'),  Y = 'Y', data = dat, 
            expand.degree = 3, cv = TRUE, group.exact = c(1, 1, 0, 0, 0))
summary(out)
```

### User-Supplied Penalties
If researchers have prior knowledge about covariates and would like to supply specific penalties for each group of covariates, we can do so by passing a vector of penalties to the `group.alpha` argument. For example, we can manually set penalties to be 0 for the the linear terms and two-way interactions and 100 for other groups:
```{r}
out <- hbal(Treat = 'D', X = c('X1', 'X2', 'X3'),  Y = 'Y', data = dat, 
            expand.degree = 3, group.alpha = c(0, 0, 100, 100, 100, 100))
summary(out)
```

### Controlling Serial Expansion

Sometimes we may not want to perform serial expansion on all covariates but instead expand on a select set of covariates, we can do so by using the `X.expand` argument to specify the covariates we want to expand on. For example, we can expand only on **X1**:
```{r}
out <- hbal(Treat = 'D', X = c('X1', 'X2', 'X3'),  Y = 'Y', data = dat, 
            expand.degree = 3, X.expand = c('X1', 'X2'))
summary(out)
```

### Selecting Covariates
Performing serial expansion on many covariates can result in a prohibitive number of covariates that need to be balanced on. In such cases, users many want to reduce the number of covariates by using the double selection method by @belloni2014inference. This screens the expanded covariates and only keeps those that are predictive for the treatment assignment or the outcome. Users can enable double selection by setting `ds = TRUE`. In the following case, almost all the higher-order terms except `X1*X2` and `X2*X3` are screened out by the double selection procedure.
```{r}
out <- hbal(Treat = 'D', X = c('X1', 'X2', 'X3'),  Y = 'Y', data = dat, 
            expand.degree = 3, ds = TRUE) 
summary(out)
att(out)
```

### Keeping/Excluding Covariates
If there are covariates that users would like to keep in the balancing conditions regardless of whether they are selected in double selection, we can use the `X.keep` argument to specify the covariates to keep. On the other hand, if a priori we know some combinations of the covariates are nonsensical, we can exclude them explicitly by using the `exclude` argument. For example, we can exclude any interaction that involves **X1** and **X2**:
```{r}
out <- hbal(Treat = 'D', X = c('X1', 'X2', 'X3'),  Y = 'Y', data = dat, 
            expand.degree = 3, exclude = list(c('X1', 'X2')))
summary(out) # X1.X2 and X1.X1.X2 removed from balancing scheme
att(out) 
```

## Example 1: Lalonde Data 

To show the use of **habl** in practice, here we use **habl** on a dataset that contains the subset of the @lalonde1986evaluating dataset from @dehejia1999causal and the Panel Study of Income Dynamics (PSID-1), which is also shipped with **hbal**. 
```{r}
data(hbal)
head(lalonde)
```

First, we adjust for the linear terms only:
```{r}
xvars <- c("age","black","educ","hisp","married","re74","re75","nodegr","u74","u75") # covariates
# hbal w/ level terms only
hbal.out <- hbal(Treat = 'nsw', X = xvars,  Y = 're78', data = lalonde) 
summary(hbal.out)
att(hbal.out)
```

Next, we add higher-order terms. The confidence intervals of both estimates cover the experimental benchmark (~$1800). As a result of cross-validation, no penalty is imposed on the linear and squared terms.  Note that `exclude=list(c("educ", "nodegr"))` removes the nonsensical interaction between `educ` and `nodegr`.
```{r}
hbal.full.out <- hbal(Treat = 'nsw', X = xvars, Y = 're78', data = lalonde, 
                      expand.degree = 2, cv = TRUE, exclude=list(c("educ", "nodegr")))
summary(hbal.full.out)
att(hbal.full.out)
```

We can check the penalties applied to each group and the cavariate balance before and after balancing.
```{r, fig.height = 8, fig.align = "left"}
hbal.full.out$group.penalty
plot(hbal.full.out)
```

## Example 2: Black and Owens (2016)

The second example comes from @black2016courting, in which the authors study the effect of promotion prospect to the Supreme Court on the behavior of circuit court judges. Here we focus on whether circuit court judges who are on the shortlist to fill Supreme Court vacancies ("contenders") ruled in line with the president as the outcome of interest. We load the dataset `contenderJudges`, which is shipped with **hbal**.
```{r}
data(hbal)
```

First, we take a look at the data structure. The outcome variable is `presIdeoVote` and the treatment variable is `treatFinal0`, indicating whether there was a Supreme Court vacancy at the time. There are also 7 covariates on judge and court characteristics and a variable `judge` that indicates the judges' names.
```{r}
str(contenderJudges)
```

We can estimate the effect of Supreme Court vacancy on judges' rulings while controlling for functions of the covariates (to the second order). We see that contender judges are more likely to rule in line with the ideology of the sitting president during a Supreme Court vacancy.

```{r}
xvars <- c("judgeJCS", "presDist", "panelDistJCS", "circmed", "sctmed", "coarevtc", "casepub")
out <- hbal(Treat = 'treatFinal0', X = xvars, Y = 'presIdeoVote', data = contenderJudges,
            expand.degree = 2, cv = TRUE)
summary(out)
att(out)
```

We can further check covariate balance before and after balancing by checking the balance plots. Here we see that the linear terms are exactly balanced between the treatment and the control groups. Imbalance among higher-order terms and interactions are also significantly reduced.

```{r, fig.height = 8, fig.align = "left"}
plot(out)
```

# Reference


